---
title: "Take-home Exercise 3: GIS-based Network Accessibility Analysis"
description: |
  This handout provides the context, the task, the expectation and the grading criteria of Take-home Exercise 3.  Students must review and understand them before getting started with the take-home exercise. 
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Setting the scene

Geographical accessibility is an important issue for the sustainable development of cities.  In this take-home exercise, you are required to apply GIS-based network accessibility analysis approach to assess accessibility to community facilities in Singapore by using publicly available data.  

## The Tasks

The specific tasks of the study are as follow:

+ derive a hexagon layer (500m by 500m) from Planning Subzone layer.  The hexagon layer should exclude all the outer island including Pulau Ubin and Pulau Tekong.,
+ compute nearest distance from the centroid of each hexagon to the following social services: 
  + eldercare (driving),
  + CHAS clinic (driving), 
  + primary school (driving).
  + childcare (walking), 
  + kindergarten (walking),   
+ using appropriate thematic mapping technique, prepare a series of thematic maps showing accessibility of  the community services above,
+ using appropriate statistical method(s) provided by QGIS, prepare a composite accessibility layer by combining these five layers,
+ plot appropriate statistics graphics to complement the GIS maps analysis, and
+ Critically discuss the geographical patterns revealed by the maps and statistical graphs. 

## Take-home Exercise Artefact

### GIS data repository

The GIS repository includes but not limited to geospatial data compiled and derived, QGIS project file and data dictionary. It must be in a single zipped file (i.e. .zip).  The geospatial data must be stored in a GeoPackage database format.  The data dictionary can be in either MS Word document or edited into the GIS data.The project artefact must be uploaded onto eLearn.

### Take-home Exercise Report

You are required to edit your take-home exercise report in MS Word format.  The take-home exercise report, beside others, should include all the thematic and choropleth maps prepared and their respective discussion.  

More importantly, the report must provide a **reproducible** step-by-step guide on the following process:

+ data compilation, extraction and integration,
+ data cleaning, preparation and wrangling,
+ GIS analysis (including tabular and graphical analysis), and
+ GIS maps design.

>Note: **Reproducible** means that readers are able to perform the same analysis and obtain similar results by using the same data sets and by following the step-by-step guide. 

The title of the report should be in the form of *SMT201_AY2021-22T1_Ex3*.

>Note: This is an individual exercise. You are required to work on the take home exercise and prepare submission individually.

## Grading

+ Quality of the GIS data model built (including metadata) (20 marks),
+ Appropriateness of the GIS methods used (20 marks),  
+ Quality of GIS maps prepared (20 marks),
+ Reproducibility of the GIS processes (20 marks), and
+ Ability to provide correct interpretation and discussion of the analysis results (20 marks).

## Submission Date

The take-home exercise artefact must be uploaded onto eLearn by the submission deadline stated below. 

Due Date: 7th November 2021 (Sunday), 11:59pm (mid-night). 


```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("_site/Take-home_Ex03.html")
```
